Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2025-04-06 manuscript/manuscript.qmd
Revised abstract

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
# author-note:
  # disclosures:
    # gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Advice that improves decision-making does not always feel good. While advice has the potential to improve decision-making by helping advisees more accurately assess feasibility and predict potential outcomes, it can also lead to disappointment when it challenges the advisee’s overly optimistic expectations. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In study 1, advisers use flattering advice to create more positive impressions, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate. Even advisers incentivised for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in financial loss for them and the advisees (Study 2). The concern for interpersonal consequences may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

keywords: [Advice, Interpresonal relationship, Gender difference]
bibliography: bibliography.bib
floatsintext: true
mask: false
link-citations: false
format:
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    prefer-html: true
  apaquarto-docx:
    default: true
    prefer-html: true
editor: 
  markdown: 
    wrap: 72
---

>>>>

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
# author-note:
  # disclosures:
    # gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Advice can improve decisions when it offers an unbiased evaluation of information, but the painful truth can disappoint an advisee's overoptimistic expectations. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers avoid such disappointment at the expense of accuracy and their own earnings. Participants inflate the attractiveness of advisees and do so to a greater extent when incentivized to be viewed as likeable. Advisers who provide flattering advice are rated as more likable and trustworthy, even as their advice is less honest and accurate (Study 1). When advisers observe an advisee's quiz performance along with a signal expressing a preference for competing against a group of low or higher performers, the (uninformative) signal influences their advice (Study 2). Finally, in a context where men expect to do better than women, advisers who observe expectations (but not gender) on top of actual performance are more likely to tell men to enter a tougher (but potentially more rewarding) competition. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice.

keywords: [advice, belief-based utility, interpersonal relationships, gender, discrimination]
bibliography: bibliography.bib
floatsintext: true
mask: false
link-citations: false
format:
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    prefer-html: true
  apaquarto-docx:
    default: true
    prefer-html: true
editor: 
  markdown: 
    wrap: 72
---
